---
title: "Sample Size Justification"
output: html_document

date: '`r format(Sys.time(), "%B %d, %Y")`'

params:
  description_population: NA,
  collect_entire_population: NA,
  description_constraints: NA,
  effect_of_interest: NA,
  minimal_detectable: NA,
  expected_effect: NA,
  expected_effect_metric_meta: NA,
  expected_effect_metric_study: NA,
  expected_effect_value_meta: NA,
  expected_effect_value_study: NA,
  expected_effect_from_meta: NA, 
  sesoi_effect_value: NA,
  sesoi_effect_metric: NA,
  statistically_detectable_effect_value: NA,
  statistically_detectable_effect_metric: NA,

  describe_similarity_meta: NA,
  describe_homogeneity_meta: NA,
  describe_bias_meta: NA,
  expected_effect_from_study: NA,
  describe_similarity_study: NA,
  describe_uncertainty_study: NA,
  describe_bias_study: NA,
  width_effect_estimate: NA,
  sensitivity_power: NA,
  distribution_effect: NA,
  meta_analysis: NA,
  will_meta_be_performed: NA,
  decision: NA,
  relative_cost: NA,
  alpha_level: NA,
  power: NA,
  effect_size: NA,
  effect_metric_1: NA,
  effect_metric_2: NA,
  relative_cost_code: NA,
  estimate: NA,
  interval_metric: NA,
  desired_accuracy: NA,
  assurance: NA,
  estimation_code: NA,
  power_goal: NA,
  power_type: NA,
  power_analysis_code: NA,
  heuristic: NA,
  heuristic_details: NA,
  justification: NA,
  no_justification_details: NA,
  observations: NA,
  participants: NA,
  participants_details: NA,
  informational_value: NA,
  
---


```{r setup, warning=FALSE, message=FALSE, echo=FALSE}
#options(knitr.table.format = 'markdown')
```


Below are four sections of the sample size justification. Part A contains a description of the population, as well as a description of the resource constraints that determine how much of the population can be sampled. In part B a description of which effect sizes are of interest is provided. In part C an overview of the inferential goal of the study is specified. In part D the sample size that will be collected is reported, and the informational value of the study is evaluated. 

# A: Sample Description

**Description of the population**

```{r, echo = FALSE, comment = NA}
cat(gsub(pattern = "\n", replacement = "  \n", x = as.character(params$description_population)))
```

**Can you collect data from the entire population?**

```{r, echo = FALSE, comment = NA}
cat(params$collect_entire_population)
```

**Description of resource constraints**

```{r, echo = FALSE, comment = NA}
cat(gsub(pattern = "\n", replacement = "  \n", x = as.character(params$description_constraints)))
```
`r if(params$description_constraints == ""){paste("*No further details about the resource constraints were provided*")}`

# B: Effects of Interest

`r if(params$effect_of_interest != "" | (is.na(params$sesoi_effect_value) == FALSE)){paste("**Information about the Smallest Effect Size of Interest**:")}`

`r if(is.na(params$sesoi_effect_value) == FALSE){paste0("The smallest effect size of interest size is specified as a ", params$sesoi_effect_metric, " of ", params$sesoi_effect_value,".")}`

`r if(params$effect_of_interest != ""){paste("The following details were provided about the smallest effect size of interest:")}`

```{r echo = FALSE, comment = NA}
if(params$effect_of_interest != ""){cat(gsub(pattern = "\n", replacement = "  \n", x = params$effect_of_interest))}
```

`r if(params$minimal_detectable != "" | (is.na(params$statistically_detectable_effect_value) == FALSE)){paste("**Information about the Minimal Statistically Detectable Effect**:")}`

`r if(is.na(params$statistically_detectable_effect_value) == FALSE){paste0("The minimal statistically detectable effect is specified as a ", params$statistically_detectable_effect_metric, " of ", params$statistically_detectable_effect_value,".")}`

`r if(params$minimal_detectable != ""){paste("The following details were provided about the minimal statistically detectable effect:")}`

```{r echo = FALSE, comment = NA}
if(params$minimal_detectable != ""){cat(gsub(pattern = "\n", replacement = "  \n", x = params$minimal_detectable))}
```


```{r echo = FALSE}
if(params$expected_effect_from_meta == "yes") {
    expected_effect_summary <- paste0("The expected effect size is based on a **meta-analysis**. The expected effect size is a ", params$expected_effect_metric_meta, " of ", params$expected_effect_value_meta,".")
    } else if (
      params$expected_effect_from_study == "yes") {
      expected_effect_summary <- paste0("The expected effect size is based on a **previous study**. The expected effect size is a ", params$expected_effect_metric_study, " of ", params$expected_effect_value_study,".")
    }
``` 

`r if(params$expected_effect_from_meta == "yes"){paste("**Information about the Expected Effect Size**")}`
`r if(params$expected_effect_from_study == "yes"){paste("**Information about the Expected Effect Size**")}`

`r if(params$expected_effect_from_meta == "yes"){paste(expected_effect_summary)}`
`r if(params$expected_effect_from_study == "yes"){paste(expected_effect_summary)}`

`r if(params$expected_effect_from_meta == "yes"){paste("An evaluation of the similarity of the meta-analysis with the planned study, a citation of the meta-analysis, and details about the effect size in the meta-analysis are provided below:")}`

```{r echo = FALSE, comment = NA}
if(params$expected_effect_from_meta == "yes"){cat(gsub(pattern = "\n", replacement = "  \n", x = params$describe_similarity_meta))}
```

`r if(params$expected_effect_from_meta == "yes"){paste("An evaluation of the heterogeneity in the meta-analysis, and which subsample of studies would be most relevant to the study that is being planned, is provided below:")}`

```{r echo = FALSE, comment = NA}
if(params$expected_effect_from_meta == "yes"){cat(gsub(pattern = "\n", replacement = "  \n", x = params$describe_homogeneity_meta))}
```

`r if(params$expected_effect_from_meta == "yes"){paste("An evaluation of whether the meta-analytic effect size estimate is unbiased, and if not, a computed effect size estimate that attempts to correct for bias, or a more conservative effect size estimate, is provided below:")}`

```{r echo = FALSE, comment = NA}
if(params$expected_effect_from_meta == "yes"){cat(gsub(pattern = "\n", replacement = "  \n", x = params$describe_bias_meta))}
```

`r if(params$expected_effect_from_study == "yes"){paste("An evaluation of the similarity of the previous study with the planned study, a citation of the previous study, and details about the effect size from the previous study are provided below:")}`

```{r echo = FALSE, comment = NA}
if(params$expected_effect_from_study == "yes"){cat(gsub(pattern = "\n", replacement = "  \n", x = params$describe_similarity_study))}
```

`r if(params$expected_effect_from_study == "yes"){paste("An evaluation of the uncertainty in the effect size estimate in the previous study, and how this is dealt with (e.g., choosing a more consevative estimate) is provided below:")}`

```{r echo = FALSE, comment = NA}
if(params$expected_effect_from_study == "yes"){cat(gsub(pattern = "\n", replacement = "  \n", x = params$describe_uncertainty_study))}
```

`r if(params$expected_effect_from_study == "yes"){paste("An evaluation of whether the effect size estimate of the previous is unbiased, and if not, any approach to correct for bias, or decisions about the use of a more conservative effect size estimate is provided below:")}`

```{r echo = FALSE, comment = NA}
if(params$expected_effect_from_study == "yes"){cat(gsub(pattern = "\n", replacement = "  \n", x = params$describe_bias_study))}
```

`r if(params$expected_effect != ""){paste("The following additional information about the expected effect size has been provided:")}`

```{r echo = FALSE, comment = NA}
if(params$expected_effect != ""){cat(gsub(pattern = "\n", replacement = "  \n", x = params$expected_effect))}
```

`r if(params$width_effect_estimate != ""){paste("The following information about the **Expected Width of the Confidence Interval** has been provided:")}`

```{r echo = FALSE, comment = NA}
if(params$width_effect_estimate != ""){cat(gsub(pattern = "\n", replacement = "  \n", x = params$width_effect_estimate))}
```

`r if(params$sensitivity_power != ""){paste("The following information about the **Sensitivity Power Analysis** has been provided:")}`

```{r echo = FALSE, comment = NA}
if(params$sensitivity_power != ""){cat(gsub(pattern = "\n", replacement = "  \n", x = params$sensitivity_power))}
```

`r if(params$distribution_effect != ""){paste("The following information about the **Distribution of Effect Sizes in This Research Area** has been provided:")}`

```{r echo = FALSE, comment = NA}
if(params$distribution_effect != ""){cat(gsub(pattern = "\n", replacement = "  \n", x = params$distribution_effect))}
```


# C: Inferential Goal

`r if(params$meta_analysis == "yes"){paste("The goal of the study is to collect data for a **future meta-analysis**. The following information about the probability that a future meta-analysis will be performed has been provided:")}`

`r if(params$meta_analysis == "yes"){paste(params$will_meta_be_performed)}`

`r if(params$meta_analysis == "yes" && params$will_meta_be_performed == ""){paste("*No further details about the probability that a meta-analysis will be performed were provided*")}`


```{r echo = FALSE}

power_metric_1 <- ifelse(params$effect_metric_1 != "Other. Specify in the Sample Size Computation field", params$effect_metric_1, "")

power_metric_2 <- ifelse(params$effect_metric_2 != "Other. Specify in the Sample Size Computation field", params$effect_metric_2, "")

estimate_metric <- ifelse(params$interval_metric != "Other. Specify in the Sample Size Computation field", as.character(params$interval_metric), "Interval (see below for details)")

assurance <- ifelse(params$assurance, paste0(" with an assurance level of ", params$assurance), "")


if(params$power_type == "a-priori power analysis"){
  power_desired <- paste0(", and the desired power is ", params$power,".")
}
if(params$power_type == "compromise power analysis"){
  power_desired <- paste0(", and the desired power is ", params$power,".")
}
if(params$power_type == "sensitivity power analysis"){
  power_desired <- paste("")
}
  
if(params$power_goal == "yes") {
    inferential_goal_summary_power <- paste0("The **inferential goal** is to perform a hypothesis test with a certain **statistical power**, computed by a ", params$power_type, ". The chosen alpha level is ", params$alpha_level, power_desired, " A justification for the chosen alpha level and desired power (or for a sensitivity power analysis, the achieved power for effects of interest), and details of the power calculation (preferably in reproducible code) is provided below.")
    } 

if(params$estimate == "yes") {
      inferential_goal_summary_accuracy <- paste0("The **inferentional goal** to estimate parameters with a desired **accuracy** based on a ", params$desired_accuracy, " ", estimate_metric, ". Details related to the sample size computation:")
    } 
if (params$decision == "yes") {
      inferential_goal_summary_decision <- paste0("The **inferentional goal** is to make a **decision**. It has been decided that making a Type I error is ", params$relative_cost, " times as costly as a Type II error. The chosen alpha level is ", params$alpha_level, ", the desired power is ", params$power,", which is based on an effect size ", power_metric_1, " of ", params$effect_size,". Details related to the sample size computation:")
    }

if(params$power_goal == "yes") {
    code_summary_power <- params$power_analysis_code
} else if(params$power_goal == "no") {
  code_summary_power <- ""
}

if (params$estimate == "yes") {
      code_summary_accuracy <- params$estimation_code
}  else if(params$estimate == "no") {
  code_summary_accuracy <- ""
}
if (params$decision == "yes") {
      code_summary_decision <- params$relative_cost_code
}  else if(params$decision == "no") {
  code_summary_decision <- ""
}
```

`r if(params$power_goal == "yes"){paste("The following information about the inferential goal related to **statistical power** has been provided:")}`

`r if(params$power_goal == "yes"){paste(inferential_goal_summary_power)}`

`r if(params$power_goal == "yes" && code_summary_power == ""){paste("*No further details or justifications for chosen values were provided*")}`

```{r echo = FALSE, comment = NA}
if(code_summary_power != ""){
  cat(gsub(pattern = "\n", replacement = "  \n", x = code_summary_power))
}
```

`r if(params$estimate == "yes"){paste("The following information about the inferential goal related to **estimation** has been provided:")}`

`r if(params$estimate == "yes"){paste(inferential_goal_summary_accuracy)}`

`r if(params$estimation == "yes" && code_summary_accuracy == ""){paste("*No further details or justifications for chosen values were provided*")}`

```{r echo = FALSE, comment = NA}
if(code_summary_accuracy != ""){
  cat(gsub(pattern = "\n", replacement = "  \n", x = code_summary_accuracy))
}
```

`r if(params$decision == "yes"){paste("The following information about the inferential goal related to **a decision** has been provided:")}`

`r if(params$decision == "yes"){paste(inferential_goal_summary_decision)}`

`r if(params$decision == "yes" && code_summary_decision == ""){paste("*No further details or justifications for chosen values were provided*")}`

```{r echo = FALSE, comment = NA}
if(code_summary_decision != ""){
  cat(gsub(pattern = "\n", replacement = "  \n", x = code_summary_decision))
}
```

`r if(params$heuristic == "yes"){paste("The sample size was not determined in relation to a specific inferential goal, but based on heuristics, about which the following information has been provided:")}`

`r if(params$heuristic == "yes" && params$heuristic_details != ""){paste(params$heuristic_details)}`

`r if(params$heuristic == "yes" && params$heuristic_details == ""){paste("*No further details or justifications for chosen values were provided*")}`

`r if(params$justification == "yes"){paste("The sample size was not determined in relation to a specific inferential goal, and no justification was given, about which the following information has been provided:")}`

`r if(params$justification == "yes" && params$no_justification_details != ""){paste(params$no_justification_details)}`

`r if(params$justification == "yes" && params$no_justification_details == ""){paste("*No further details were provided*")}`

# D: Informational Value of the Study

Based on the resource constraints, the effects of interest, and the inferential goals, the following evaluation of the informational value of the study has been provided:

Given the following resource constraints: 

```{r, echo = FALSE, comment = NA}
cat(gsub(pattern = "\n", replacement = "  \n", x = as.character(params$description_constraints)))
```

```{r echo = FALSE}
# Final summary
if(params$expected_effect_from_meta == "yes") {
    expected_effect_final_summary <- paste0("an expected effect size of ", params$expected_effect_metric_meta, " = ", params$expected_effect_value_meta,"")
    } else if (
      params$expected_effect_from_study == "yes") {
      expected_effect_final_summary <- paste0("an expected effect size of ", params$expected_effect_metric_study, " = ", params$expected_effect_value_study,"")
    }
``` 

and given `r if(is.na(params$sesoi_effect_value) == FALSE){paste0("a smallest effect size of interest size of ", params$sesoi_effect_metric, " = ", params$sesoi_effect_value,",")}` `r if(is.na(params$statistically_detectable_effect_value) == FALSE){paste0("a minimal statistically detectable effect of ", params$statistically_detectable_effect_metric, " = ", params$statistically_detectable_effect_value,",")}` `r if(params$expected_effect_from_meta == "yes"){paste0(expected_effect_final_summary,",")}` `r if(params$expected_effect_from_study == "yes"){paste0(expected_effect_final_summary,",")}` `r if(params$width_effect_estimate != ""){paste0("an evaluation of the expected width of a confidence interval as described below:")}`

```{r echo = FALSE, comment = NA}
if(params$width_effect_estimate != ""){cat(gsub(pattern = "\n", replacement = "  \n", x = params$width_effect_estimate))}
```

`r if(params$sensitivity_power != ""){paste0("an evaluation of effects one has sufficient power to detect based on a sensitivity power analysis as described below:")}`

```{r echo = FALSE, comment = NA}
if(params$sensitivity_power != ""){cat(gsub(pattern = "\n", replacement = "  \n", x = params$sensitivity_power))}
```

`r if(params$distribution_effect != ""){paste0("and a distribution of effect sizes in the literature as described below:")}`

```{r echo = FALSE, comment = NA}
if(params$distribution_effect != ""){cat(gsub(pattern = "\n", replacement = "  \n", x = params$distribution_effect))}
```

```{r echo = FALSE}

if(params$meta_analysis == "yes") {
    inferential_goal_final_meta <- paste0("given the inferential goal is to collect data for a future meta-analysis, ")
} else if (params$meta_analysis == "no") {
      inferential_goal_final_meta <- ""
    }
if(params$power_goal == "yes") {
    inferential_goal_final_power <- paste0("given the inferential goal based on a ", params$power_type, " with an alpha level of ", params$alpha_level, power_desired,", ")
} else if (params$power_goal == "no") {
      inferential_goal_final_power <- ""
    } 
if (params$estimate == "yes") {
      inferential_goal_final_estimate <- paste0("given the inferential goal to estimate parameters with a desired accuracy ", params$desired_accuracy, " ", estimate_metric,", ")
} else if (params$estimate == "no") {
      inferential_goal_final_estimate <- ""
    }
if (params$decision == "yes") {
      inferential_goal_final_decision <- paste0("given the inferential goal to make a decision based on the judgment that a Type I error is ", params$relative_cost, " times as costly as a Type II error, the chosen alpha level is ", params$alpha_level, ", and the desired power is ", params$power,", ")
} else if (params$decision == "no") {
      inferential_goal_final_decision <- ""
    }
if (params$heuristic == "yes") {
      inferential_goal_final_heuristic <- paste0("given the inferential goal was not specified, but a heuristic was used instead, ")
} else if (params$heuristic == "no") {
      inferential_goal_final_heuristic <- ""
    }
if (params$justification == "yes") {
      inferential_goal_final_justification <- paste0("given the inferential goal was not specified, and no justification for the sample size was provided, ")
} else if (params$justification == "no") {
      inferential_goal_final_justification <- ""
    }

```

and `r paste(inferential_goal_final_power)``r paste(inferential_goal_final_estimate)``r paste(inferential_goal_final_decision)``r paste(inferential_goal_final_heuristic)``r paste(inferential_goal_final_justification)``r paste(inferential_goal_final_meta)`

the **sample size** in the planned study will consist a total of `r params$participants` participants, each contributing `r params$observations` observations. 

The following additional details about the sample size were provided: 

```{r echo = FALSE, comment = NA}
if(params$participants_details != ""){cat(gsub(pattern = "\n", replacement = "  \n", x = params$participants_details))}
```

`r if(params$participants_details == ""){paste("*No further details about the sample size were provided*")}`

An explanation of the informational value of the sample size that will be collected, given any resource constraints, the effects of interest, and the inferential goal, is provided below:

```{r echo = FALSE, comment = NA}
cat(gsub(pattern = "\n", replacement = "  \n", x = params$informational_value))
```

